---
title: "R Day 2"
author: "Shreyas Sabnis"
date: "9/18/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
here::here("data", "gapminder5.csv")

gapminder <- read.csv(here::here("data","gapminder5.csv"))

str(gapminder)

```

## Converting factors to characters

```{r}
gapminder$country=as.character(gapminder$country)
gapminder$continent=as.character(gapminder$continent)
str(gapminder)
```

```{r}
# create a vector of values that you want to repeat the function for 
#obs <- 1:nrow(gapminder)
#initialize the for loop with 'for (i in vector)'
#for (i in obs) {
#  gapminder[i,'gdp'] <- gapminder[i, 'pop'] * gapminder[i, 'gdpPercap]
#}
```

##New variable that get the natural log of the GDP per cap and population
```{r}
#obs <- 1:nrow(gapminder)
#for (i in obs) {
#  gapminder[i,'log_gdpPercap'] = log(gapminder[i, 'gdpPercap'])
#  gapminder[i,'log_pop'] = log(gapminder[i, 'pop'])
#}
```


##vectorized function
```{r}
gapminder$vec_log_gdpPercap = log(gapminder$gdpPercap)
all(gapminder$vec_log_gdpPercap == gapminder$log_gdpPercap)
```

##has life expectancy increased over time?
```{r}
#years <- unique(gapminder$year)

#for (i in year) {
#  mean_ye <- mean(gapminder$lifeExp[gapminder$year == i], na.rm= T)
#  print(paste0(i, ":", mean_le))
#}
```

#Life expectancy by continent
```{r}
continent <- unique(gapminder$continent)

for (i in continent){
  mean_con <- mean(gapminder$lifeExp[gapminder$continent == i], na.rm= T)
  print(paste0(i, ":", mean_con))
}
```

## LE by continent and year
```{r}
conts <- unique(gapminder$continent)
years <- unique(gapminder$year)

for (i in conts){
  for (j in years) {
    print(paste0(i," ", j," : ", mean(gapminder$lifeExp[ gapminder$continent == i & gapminder$year ==j])))
  }
}
```

##LE by continent and country
```{r}
conts <- unique(gapminder$continent)
years <- unique(gapminder$year)

for (i in conts) {
  for (j in years) {
    print(paste0(i, " ", j, " ", sd(gapminder$lifeExp[ gapminder$continent == i & gapminder$year == j])))
  }
}
```

##Apply function
```{r}
apply(gapminder[, c('lifeExp','pop','gdpPercap')], 2 , mean, na.rm = TRUE)
```
#lapply
```{r}
lapply(gapminder, mean)
```

#Anonymous functions in apply
```{r}
sapply(unique(gapminder$year), function(x) mean(gapminder$lifeExp[gapminder$year== x]))
```

#While loop
#LE for years before 1987
```{r}
i= min(gapminder$year)
while (i < 1987){
  print(paste0 (i , " : ", sd(gapminder$lifeExp[gapminder$year == i])))
  i = i+ 5
}
```

## SD of LE b/w 1987 and 2002
```{r}
i <- 1987
while(i <= 2002) {
  print(paste(i, sd(gapminder$lifeExp[gapminder$year == i])))
  i = i + 5
}
```

#if else in a for loop
```{r}
thresh = 70
for (i in unique(gapminder$continent)) {
  measure = mean(gapminder$lifeExp[gapminder$continent== i])
  if (measure <70) {
    print(paste(i, "has a life exp of", measure, "which is less than 70"))
  } else {
    print(paste(i, "has a life exp of", measure, "which is not less than 70"))
  }
}
```

#mean population for years greater than or equal to 1987
```{r}
thresh = 1987

for (i in unique(gapminder$year)){
  if (i <1987){
    print(paste(i,' is less than 1987'))
  } else {
print( paste('the mean population for ', i, 'is', mean(gapminder$pop[gapminder$year == i])))
  }
}
```

#Functions
```{r}
get_vals <- 
  function(df, variable = 'continent') {
  print(paste(variable, ":", unique(df[[variable]])))
  }

get_vals(gapminder,'continent')
```


##mean and standard deviation for life expentancy for a given country in the gapminder dataset


##Write a function that reports the mean, median, minimum, and maximum for life expectancy for a continent in gapminder
```{r}
get_summary_stats <-
  function(df,variable,continent) {
    mean_val = mean(df[[variable]][df$continent == continent])
    med_val = median(df[[variable]][df$continent == continent])
    min_val = min(df[[variable]][df$continent == continent])
    max_val = max(df[[variable]][df$continent == continent])
    print(paste(" summary statistics for ",continent,": (",mean_val,med_val,min_val,max_val,")"))
  }
get_summary_stats(gapminder, 'pop', 'Asia')
```

```{r}
get_summary_stats(gapminder, 'lifeExp', 'Asia')

```

##Loglog model, life exp vs GDP
```{r}
viz <-
  function(df,dx,dy,year) {
    dat = df[df$year== year,]
    y = log(dat[[dy]])
    x= log(dat[[dx]])
    fit <- lm(y~x)
    plot(y~x, main = year, xlab= dx, ylab= y)
    lines(x, predict(fit), col='blue')
  }
```

```{r}
viz(gapminder, 'pop', 'gdpPercap', 1977)
```


